---
title: "p8105_hw3_mrc2229"
author: "Matthew Curran"
date: "October 12, 2019"
output: github_document
---

```{r setup, include=FALSE}
library (tidyverse)
library(viridis)
library (leaflet)
library(knitr)
```

#Problem 1
##Bring in instacart data
```{r}
library(p8105.datasets)
data("instacart")
```
##Examining the instacart data
There are `r nrow (instacart)` observations in the instacart dataset and `r ncol (instacart)` variables in the instacart data set. Some of the key variables are product_name, which shows the product purchased; aisle, which tells us the aisle the purchased item comes from; and order_id, which tells us the id of the person who ordered a certain item. So we can see things like that person with order_id 1 got  `r instacart %>% filter (order_id==1, aisle_id==24) %>% summarise(n_distinct(product_name))` items from aisle 24, the fresh fruit aisle. The instacart dataset also gives information on the time an order was placed. As an example, the person with order_id 1 made their order at the `r instacart %>% filter (order_id==1) %>% summarise(max(order_hour_of_day))`th hour of the day.

##Examining the aisles
There are `r instacart %>% summarise(n_distinct(aisle))` aisles in total and the most items are ordered from the `r instacart %>% count(aisle,name="aisle_count") %>% mutate(aisle_count_rank = rank(desc(aisle_count))) %>% filter (aisle_count_rank == 1 ) %>% summarise(aisle)` and `r instacart %>% count(aisle,name="aisle_count") %>% mutate(aisle_count_rank = rank(desc(aisle_count))) %>% filter (aisle_count_rank == 2 ) %>% summarise(aisle)` aisles.

##Plot of number of items ordered in each aisle
```{r}
items_per_aisle_plot = instacart %>%
  count(aisle,name="aisle_count") %>%
  filter(aisle_count>10000) %>%
  arrange (desc(aisle)) %>%
  ggplot(aes(x=aisle, y=aisle_count)) +
  geom_bar(stat="identity", fill="blue") +
  labs(
    title = "Ammount of Items Purchased in Each Aisle",
    x = "Aisle",
    y ="Ammount of Items Purchased"
  )

items_per_aisle_plot + coord_flip()
```

##Table of most popular items in "baking ingrediant", "dog food care", and "packaged vegatables fruits"
```{r}
most_popular_items = instacart %>%
  select(aisle,product_name) %>%
  filter(
    aisle == "baking ingredients"|
    aisle == "packaged vegetables fruits"|
    aisle == "dog food care"
    ) %>%
  group_by(aisle) %>%
  count(product_name,name="product_count")%>%
  mutate(product_count_rank = rank(desc(product_count))) %>% 
  filter (product_count_rank < 4) %>%
  arrange(aisle, product_count_rank) %>%
  select(aisle, product_count_rank, product_name, product_count) %>%
  kable()
  
most_popular_items
```

##Table showing day and time Pink Lady Apples and Coffee Ice Cream are ordered
```{r}
pink_lady_coffee = instacart %>%
  filter(
  product_name == "Pink Lady Apples"|
  product_name == "Coffee Ice Cream"
  ) %>%
  arrange (product_name, order_dow) %>%
  mutate(
    order_dow = 
    recode(order_dow,
    "0" = "Sunday",
    "1" = "Monday",
    "2" = "Tuesday",
    "3" = "Wednesday",
    "4" = "Thursday",
    "5" = "Friday",
    "6" = "Saturday")
    ) %>%
   mutate(order_dow = fct_relevel(order_dow, "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")) %>%
  group_by(product_name,order_dow) %>%
  summarise(mean_hour_of_day=mean(order_hour_of_day)) %>%
  separate(mean_hour_of_day, into = c("hour", "minutes"), sep = 2) %>%
  mutate(minutes = as.numeric(minutes),
         minutes = round(minutes*60, digits = 0),
         hour = as.numeric(hour),
         mean_time_of_day = paste(hour,minutes,sep = ":")) %>%
  select(product_name, order_dow, mean_time_of_day) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_time_of_day
  ) %>%
  rename(Product=product_name) %>%
  kable()

pink_lady_coffee
```

#Problem 2
##Bring in and clean BRFSS data
```{r}
library(p8105.datasets)
data("brfss_smart2010")
brfss_data = brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health" |
  response == "Poor"|
  response == "Fair"|
  response == "Good"|
  response == "Very good"|
  response == "Excellent"
  ) %>%
  mutate(response = fct_relevel(response, "Poor", "Fair", "Good", "Very good", "Excellent"))
```

##States with 7 or more locations
```{r}
state_locations = brfss_data %>%
  filter(year == 2002 | year ==2010) %>%
  group_by(year, locationabbr) %>%
  summarise(number_of_state_locations = n_distinct(locationdesc)) %>%
  filter (number_of_state_locations >= 7)
```

In `r state_locations %>% filter (year==2002) %>% summarise(n_distinct(locationabbr))` states had 7 or more locations.Those states were CT, FL, MA, NC, NJ, and PA.

In `r state_locations %>% filter (year==2010) %>% summarise(n_distinct(locationabbr))` states had 7 or more locations.Those states were CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA.

##An "Excellent" response only dataset and speghetti plot.
```{r}
excellent_data = brfss_data %>%
  filter (response == "Excellent") %>%
  group_by(year,locationdesc) %>%
  mutate(ave_data_value = mean(data_value, na.rm = TRUE)) %>%
  ungroup(locationdesc) %>%
  select(year, locationabbr, ave_data_value) %>%
  distinct() %>%
  drop_na()%>%
  ggplot(aes(x=year,y=ave_data_value))+
  geom_line(aes(group=locationabbr, color=locationabbr))+
  scale_color_viridis(
    discrete = TRUE,
    name = "State Color Key"
  )+
  labs(
    title = "Average Data Value over Time within a State",
    x = "Year",
    y ="Average Data Value (%)"
  ) 
  
excellent_data
```

##NY State distribution of data values for each responce type in 2006 and 2010
```{r}
NYS_data = brfss_data %>%
  filter(year == 2006 | year == 2010) %>%
  filter (locationabbr == "NY") %>%
  ggplot(aes(x=locationdesc, y=data_value, fill=response))+
  geom_bar(stat = "identity", position = "dodge")+
  facet_grid(~year)+
  labs(
    title = "Data Values for Responses among Locations in NY State for 2006 and 2010",
    x = "NY State County",
    y = "Response Percentages"
  ) 

NYS_data + coord_flip()
```